---
title: "hw6"
author: "Cory Costello"
date: "November 7, 2017"
output: pdf_document
---
#1 Read and Tidy Data

This was a little tricky, since the spreadsheet had 7 lines at the beginning that needed to be skipped and a line at the end also need to be skipped. 

I got through this by using the `skip = 7` argument to skip the first 7 lines, then you had to set `header = TRUE` to get the header in correctly (which started on line 8), then I passed that file to `slice()` to save just the first three rows (eliminating the 3 empty rows, and the final row of notes about the data).

Finally, I gathered the data (excluding the labor status variable), so that year was represented row-wise. Then, I had to use a combination of separates and unites to get percentage and se variables together, and ended by mutating to turn the percentage, se, and year into numeric variables (year required `parse_number()`, since each had an x in front of the year, as it was imported as column names).

Finally, I used bind_rows to combine the data, used `.id = TRUE` to maintain a record of the source dataset, and then changed those id's (1, 2, 3) to substantive labels (not_happy, pretty_happy, very_happy)
```{r read_in_and_tidy_data}
library(tidyverse)
library(rio)

not_happy <- import("not_happy.csv", skip = 7, header = TRUE) %>%
  janitor::clean_names() %>%
  slice(1:3) %>%
  gather(year, pct_endorsement, -labor_force_status) %>%
  separate(pct_endorsement, c("pct_endorsement", "pct_dec",
                              "se_endorsement", "se_dec")) %>%
  unite(pct_endorsement, pct_endorsement:pct_dec, sep = ".") %>%
  unite(se_endorsement, se_endorsement:se_dec, sep = ".") %>%
  mutate(pct_endorsement = as.numeric(pct_endorsement),
         se_endorsement = as.numeric(se_endorsement),
         year = parse_number(year))
  

pretty_happy <- import("pretty_happy.csv", skip = 7, header = TRUE) %>%
  janitor::clean_names()%>%
  slice(1:3) %>%
  gather(year, pct_endorsement, -labor_force_status) %>%
  separate(pct_endorsement, c("pct_endorsement", "pct_dec",
                              "se_endorsement", "se_dec")) %>%
  unite(pct_endorsement, pct_endorsement:pct_dec, sep = ".") %>%
  unite(se_endorsement, se_endorsement:se_dec, sep = ".") %>%
  mutate(pct_endorsement = as.numeric(pct_endorsement),
         se_endorsement = as.numeric(se_endorsement),
         year = parse_number(year))
    

very_happy <- import("very_happy.csv", skip = 7, header = TRUE) %>%
  janitor::clean_names()%>%
  slice(1:3) %>%
  gather(year, pct_endorsement, -labor_force_status) %>%
  separate(pct_endorsement, c("pct_endorsement", "pct_dec",
                              "se_endorsement", "se_dec")) %>%
  unite(pct_endorsement, pct_endorsement:pct_dec, sep = ".") %>%
  unite(se_endorsement, se_endorsement:se_dec, sep = ".") %>%
  mutate(pct_endorsement = as.numeric(pct_endorsement),
         se_endorsement = as.numeric(se_endorsement),
         year = parse_number(year))

full_data <- bind_rows(not_happy, pretty_happy, very_happy, .id = "happiness") %>%
  mutate(happiness = ifelse(happiness == 1, "not_happy", 
                            ifelse(happiness == 2, "pretty_happy", "very_happy")))
```

# 2 Answering some questions

```{r 2a_pct_very_happy}
highest_pct_vhappy<- full_data %>%
  filter(happiness == "very_happy") %>%
  arrange(desc(pct_endorsement)) %>%
  slice(1)

highest_pct_vhappy
```
It looks like the highest percentage of folks that endorsed being very happy were `r highest_pct_vhappy$labor_force_status` in the year `r highest_pct_vhappy$year`; `r highest_pct_vhappy$pct_endorsement`% of them endorsed being very happy.

```{r 2b_avg_pct_happiness_employed}
pct_happiness_employed<- full_data %>%
  filter(labor_force_status == "Employed") %>%
  group_by(happiness) %>%
  summarize(m_pct_endorsement = mean(pct_endorsement, na.rm = TRUE),
            m_se_endorsement  = mean(se_endorsement, na.rm = TRUE))

pct_happiness_employed
```

It looks like, within respondents whose labor force satus was reported as Employed, `r round(pct_happiness_employed$m_pct_endorsement[1], 2)`% indicated that they were "not happy", `r round(pct_happiness_employed$m_pct_endorsement[2], 2)`% indicated that they were "pretty happy", and `r round(pct_happiness_employed$m_pct_endorsement[3], 2)`% indicated that they were "very happy", averaged across available years.

# 3 Plots!!
```{r}
library(ggthemes)
theme_set(theme_minimal())

ggplot(full_data, aes(x = year, y = pct_endorsement, color = labor_force_status)) +
  geom_line()+
  geom_smooth(se= FALSE)+
  facet_wrap(~happiness, nrow = 1)
```

```{r }
full_data %>%
  group_by (labor_force_status, happiness) %>%
  mutate(m_pct_endorsement = (mean(pct_endorsement))/100,
         pct_endorsement = pct_endorsement/100) %>%
  ggplot(aes(x = happiness, y = pct_endorsement))+
  geom_bar(aes(y = m_pct_endorsement, fill = labor_force_status), 
           stat = "identity", position = "dodge", alpha = .25)+
  geom_point(aes(y = pct_endorsement, color = labor_force_status), 
             position =  position_jitterdodge(jitter.width = .2))+
  scale_fill_brewer(palette = "Set1")+
  scale_colour_brewer(palette = "Dark2")+
  scale_y_continuous(name="Fluorescent intensity/arbitrary units", labels = scales::percent) + 
  ggtitle("Happiness by Labor Status", subtitle = "Data from 1972 to 2016 (bars represent means)")
```

I just wanted to say somewhere (so I'm putting it here) that getting the percentage sign on the y-axis was pretty tough. I ended up transforming the variables to proportions (.xx) from percentage (xx.xx) and then using the `scales::percent()` function within the `scale_y_continuous()` function. I'm curious if there is an easier way (which I guess you'll probably cover tomorrow).