---
title: "HW 2"
author: "Cory Costello"
date: "October 9, 2017"
output: pdf_document
---
# Part 1: Load Data
In addition to loading the data, I'm going to factorize treatment, setting, and sex to make the output a little prettier later.
```{r Part 1 Load data}
library(rio)
library(tidyverse)
library(knitr)
sesame13 <- import("Sesame13.sav")
sesame13$treatmen <- factorize(sesame13$treatmen)
sesame13$setting <- factorize(sesame13$setting)
sesame13$sex <- factorize (sesame13$sex)
```

# Part 2: Difference between treatments at pretest
```{r Part 2 Diff bw treatment and control}
pretest_differences <- sesame13 %>%
  select(starts_with("pre"), treatmen) %>%
  group_by (treatmen) %>%
  summarize(m_prebody  = mean(prebody, na.rm = TRUE),
            m_prelet   = mean(prelet, na.rm = TRUE),
            m_preform  = mean(preform, na.rm = TRUE),
            m_prenumb  = mean(prenumb, na.rm = TRUE),
            m_prerelat = mean(prerelat, na.rm = TRUE),
            m_preclass = mean(preclass, na.rm = TRUE))
  
kable(pretest_differences, digits = 2)
```

# Part 3
## Part 3a: Average number gain
```{r Part 3a}
number_gain <- sesame13 %>%
  select(postnumb, prenumb) %>%
  mutate(numb_gain  = postnumb - prenumb) %>%
  summarize(M_numb_gain = mean(numb_gain, na.rm = TRUE))

kable(number_gain, digits = 2)
```

Looks like the average gain in the numbers outcome was `r round(number_gain$M_numb_gain[1], 2)`. 

## Part 3b: Average number gain by treatment
```{r part 3b}
number_gain_by_treat <- sesame13 %>%
  select(treatmen, postnumb, prenumb) %>%
  mutate(numb_gain  = postnumb - prenumb) %>%
  group_by(treatmen) %>%
  summarize(m_numb_gain = mean(numb_gain, na.rm = TRUE))

kable(number_gain_by_treat, digits = 2)
```

Looks like those people who received the Encouraged Viewing treatment had a higher gain in number ability (*Mean number gain* = `r round(number_gain_by_treat$m_numb_gain[2], 2)`) than those who received no encouragment(*Mean number gain* = `r round(number_gain_by_treat$m_numb_gain[1], 2)`).

## Part 3c: Average number gain by setting
```{r Part 3c}
number_gain_by_setting <- sesame13 %>%
  select(setting, postnumb, prenumb) %>%
  mutate(numb_gain  = postnumb - prenumb) %>%
  group_by(setting) %>%
  summarize(m_numb_gain = mean(numb_gain, na.rm = TRUE))

kable(number_gain_by_setting, digits = 2)
```

Looks like people in different settings had similar gains in number ability. People in the home setting had an average gain of `r round(number_gain_by_setting$m_numb_gain[1], 2)` and people in the school setting had an average gain of `r round(number_gain_by_setting$m_numb_gain[2], 2)`.

## Part 3d: Average number gain by setting and treatment
I'm going to try to get this by just using both setting and treatment in the group by function.
```{r Part 3d}
number_gain_by_treatxsetting <- sesame13 %>%
  select(treatmen, setting, postnumb, prenumb) %>%
  mutate(numb_gain  = postnumb - prenumb) %>%
  group_by(setting, treatmen) %>%
  summarize(m_numb_gain = mean(numb_gain, na.rm = TRUE))

kable(number_gain_by_treatxsetting, digits = 2)
```

Huh, looks like that worked. I'm always amazed at how much sense functions fromt the tidyverse make.

# Part 4

## Part 4a: Gain in letter outcome by sex
```{r part 4a: Gain in letter outcome by sex}
letter_gain_by_sex <- sesame13 %>%
  select(sex, prelet, postlet) %>%
  mutate(letter_gain = postlet - prelet) %>%
  group_by(sex) %>%
  summarize (m_letter_gain = mean(letter_gain, na.rm = TRUE))

kable(letter_gain_by_sex, digits = 2)

```

It appears that average letter gain was lower for men (*mean letter gain* = `r round(letter_gain_by_sex$m_letter_gain[1], 2)`) than it was for women (*mean letter gain* = `r round(letter_gain_by_sex$m_letter_gain[2], 2)`).

## Part 4b: Most variable gain score

```{r Part 4b: Most variable gain score}
most_variable_gain_score <- sesame13 %>%
  select(setting, starts_with("pre"), starts_with("post")) %>%
  mutate(body_gain  = postbody - prebody, 
         let_gain   = postlet - prelet,
         form_gain  = postform - preform,
         numb_gain  = postnumb - prenumb,
         relat_gain = postrel - prerelat,
         class_gain = postclas - preclass) %>%
  group_by(setting) %>%
  summarize(sd_body_gain  = sd(body_gain, na.rm = TRUE),
            sd_let_gain   = sd(let_gain, na.rm = TRUE), 
            sd_form_gain  = sd(form_gain, na.rm = TRUE), 
            sd_numb_gain  = sd(numb_gain, na.rm = TRUE), 
            sd_relat_gain = sd(relat_gain, na.rm = TRUE), 
            sd_class_gain = sd(class_gain, na.rm = TRUE)) %>%
  gather(key = "gain_score", value = "sd", -setting) %>%
  arrange(desc(sd))

kable(most_variable_gain_score, digits =2)
```

As you can see, it looks like letter gain is the most variable in the School setting (*SD* = `r round(most_variable_gain_score$sd[1], 2)`) and the Home setting (*SD* = `r round(most_variable_gain_score$sd[2], 2)`).

## Part 4c: Average ppvt by treatment within site

```{r Average ppvt by treatment within site}
ppvt_by_treatment_wi_site <- sesame13 %>%
  select(treatmen, site, ppvt) %>%
  group_by (site, treatmen) %>%
  summarize (m_ppvt_score = mean(ppvt, na.rm = TRUE))

kable(ppvt_by_treatment_wi_site, digits = 2)
```

It looks like there ppvt (Peabody Picture-Vocabulary Test) scores were pretty similar within site across treatment conditions. Although I did not formally test this, it suggests that a decent amount of the variability in ppvt scores is due to site rather than treatment. It further suggests that one should use a modeling approach that accounts for the Site (probably an MLM, with students nested in site).